---
title: "Auto pre screening predictors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Auto pre screening predictors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TopK)
```

### Overview

The pre_screen_predictors function is designed to identify the most informative predictors in a dataset for a given response variable. It applies appropriate statistical tests based on the types of predictors and the response variable to determine the strength of the relationships. This vignette uses a sample dataset with various predictor types to demonstrate how the function works and which statistical tests are applied.

### Functionality

The function performs the following steps:

Pre-processing: Removes rows with missing values, issuing a warning if any NAs are found.
Statistical Tests: Applies different statistical tests depending on the type of predictor and response:
Pearson Correlation for continuous predictors with continuous responses.
T-test for binary predictors with continuous responses and continuous predictors with binary responses.
ANOVA for categorical predictors with more than two levels and continuous responses.
Chi-squared Test for categorical predictors with categorical responses.
Selection of Top Predictors: Returns the top K predictors based on the lowest p-values.

### Example Data

To illustrate the function's usage, let's create a sample dataset with various predictor types and a continuous response variable (y_continuous) and a binary response variable (y_binary).

```{r}
# Set seed for reproducibility
set.seed(123)

# Define continuous predictors
X1 <- rnorm(100, mean = 50, sd = 10)  # continuous variable
X2 <- rnorm(100, mean = 100, sd = 15)

# Define binary predictors
X3 <- sample(c("GroupA", "GroupB"), 100, replace = TRUE)  # categorical binary
X4 <- rbinom(100, 1, 0.5)  # numerical binary (0/1)

# Define categorical predictor with more than 2 levels
X5 <- sample(c("Red", "Green", "Blue"), 100, replace = TRUE)

# Define response variables
y_continuous <- rnorm(100, mean = 75, sd = 20)  # continuous response
y_binary <- sample(c("Yes", "No"), 100, replace = TRUE)  # binary response

# Create data frame with predictors
data <- data.frame(X1, X2, X3, X4, X5, y_continuous, y_binary)

```

### Applying the Function 

Using this data, let's apply the pre_screen_predictors function to identify the top 3 predictors for both the continuous and binary response variables.

#### Continuous Response Example

```{r}
# Get top 3 predictors for continuous response
result_continuous <- auto_pre_screen_predictors(data[, c("X1", "X2", "X3", "X4", "X5")], y_continuous, k = 3)

# Display results
print("Results for Continuous Response:")
print(result_continuous)

```

In this example, we have a continuous response variable (y_continuous). The function will apply Pearson Correlation for continuous predictors, ANOVA for categorical predictors, and T-test for binary predictors with continuous responses.

### Binary Response Example

```{r}
# Get top 3 predictors for binary response
result_binary <- auto_pre_screen_predictors(data[, c("X1", "X2", "X3", "X4", "X5")], y_binary, k = 3)

# Display results
print("Results for Binary Response:")
print(result_binary)

```

Here, we have a binary response variable (y_binary). The function applies Chi-squared Test for categorical predictors and T-test for continuous predictors with binary responses.

### Interpretation of Results
Given the example data, the function should return results indicating the most significant predictors based on p-values and the statistical test applied.

#### Continuous Response:
The predictors with the lowest p-values are likely to have stronger associations with the response. Pearson Correlation is used for continuous predictors like X1 and X2. ANOVA is used for categorical predictors like X5.

#### Binary Response:
For binary responses, Chi-squared Test is used for categorical predictors like X3 and X5. T-test is used for continuous predictors like X4 and X2.

### Conclusion

The pre_screen_predictors function provides a way to identify the most significant predictors in a dataset for a given response variable, using appropriate statistical tests. This vignette demonstrates its application to a sample dataset and explains the results, helping users understand which predictors are most relevant for further analysis or modeling.
